---
title: "Problem Set 2"
output:
  html_document:
    theme: spacelab
    toc: yes
    tod_depth: 1
    toc_float: yes
    df_print: paged
   
---


Ido Shalom  


Ofri Ahiel  


Daniel Rodan  


Ori Shneor  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Clear desk
rm(list=ls())

# install the ri package.
# url <- "https://cran.r-project.org/src/contrib/Archive/ri/ri_0.9.tar.gz"
# 
# install.packages(url, repos=NULL, type="source")

# Load libraries
library(psych)
library(stargazer)
library(dplyr)
library(ri)
```

```{css echo=FALSE}
th, td {
    padding-left: 20px;
    text-align: left;
}
```


# Part 1. Setup and Descriptive Statistics


## Section 1


```{r}
# Read the data
dat <- readRDS('earlytrainingproject_clean.rds')
```


## Section 2


```{r}
# set a features' vector for varsChild.
varsChild <- c("iq5", "iq6", "iq12", "retn12", "iep12")

# set a features' vector for varsTeen.
varsTeen <- c("iq15", "hsgrad", "parent")

# set a features' vector for varsAdults.
varsAdults <- c("college", "employed", "convicted", "felon", "jailed", "marijuana")

# combining the three vectors.
vars <- c(varsChild, varsTeen, varsAdults)
```


## Section 3


```{r}
# set a new data frame with the requested features.
df <- dat[c(c("SUBJECT", "DC_TRT", "SEX"), vars)]
```


## Section 4


```{r}
# Using "Describe" to display the data.
describe(df)

# Omit the NA's
df <- na.omit(df)
```


`r (count(dat) - count(df))[[1]]` observations contained NA's and were omitted.


## Section 5


```{r}
# set a binary variable "treat" that defined as 1 for the case that the observation been given treatment and 0 otherwise.
df$treat <- ifelse(df$DC_TRT == "Treatment", 1, 0)
```


## Section 6


```{r two-tables, results="asis"}
# Create two descriptive statistics tables, one for all the observtions that are in the "control" group for each variable, and the same for the "treatment" group.   
stargazer(df[df$treat == 0, vars], df[df$treat == 1, vars], type="html", title=c("Descriptive Statistics: Control", "Descriptive Statistics: Treatment"), digits=2, summary=TRUE)
```


# Part 2. Estimates of Average Treatment Effects


## Section 1


```{r results="asis"}
# Creating  a function using "tapply" that takes the mean of a given input by the category "treatment".
return_mean_by_treatment <- function(x){tapply(x, df$treat, mean)}

# Compute the means by "treatment" for each of the variables in vars.it will give us a rows of the means by "control" and "treatment" for each variable .
means <- sapply(df[vars],return_mean_by_treatment)

# calculate the difference between treatment and control means for each variable.
means <- t(rbind(means, means[2,] - means[1,]))

# defining the column names for the table.
colnames(means) <- c("Control","Treated","Difference")

# create the table using "stargazer" that will display for each variable,the estimated treatment mean,control mean and the difference between them ("the real case").
stargazer(means, type="html", title="Treatment and Control: Means at Baseline", digits=2)
```


### Section 2


```{r results='asis'}
# Running a linear regression for each variable to get the beat_0 estimator ("the control group") and the beta_1 estimator ("the estimated difference").
model <- lapply(vars, function(x) lm(paste0(x, ' ~ ', 'treat'), df))
# create the table using "stargazer" that will display for each variable his OLS Estimators.
stargazer(model, type="html", title="OLS Estimators", column.labels=vars)

```


As we can see,the beat_0 estimator is indeed the value for the "control group", and the beta_1 estimator s indeed the value for the "estimated difference".


# Part 3. Randomization Inference


## Section 1


### Sub-section a


```{r}
# computing the probability of treatment in the sample.as we expected,we got a 0.5 probability (41/82 treated observation for each variable) 
probs = genprobexact(df$treat, clustvar=df$SUBJECT, blockvar=NULL)
probs
```


### Sub-section b


```{r results='hide'}
# setting the seed to prevent multiple randomization on the same data frame, " to keep the same randomization"  
set.seed(2023)

# create a 82*5000 matrix such that, every observation get 0 or 1 by randomization, which makes the "fake" cases.
perms <- genperms(df$treat, blockvar=NULL, clustvar=df$SUBJECT, maxiter=5000)

```


## Section 2


A small clarification:
we did the parts of section 2 altogether,we thought it will be better to display the function as a whole rather then sub-section each part of it. 


```{r}
# defining the function for the "fake" distribution. 
fake_distribution<- function(x){
  # "setting the seed"
  set.seed(2023)
  # computing the probability of treatment for each of the 14 variables.
  probs <- genprobexact(df$treat, clustvar=df$SUBJECT, blockvar=NULL)
  # create a 82*5000 matrix such that, every observation in each of the 14 variables get 0 or 1 by randomization, which makes the "fake" cases.
  perms <- genperms(df$treat, clustvar=df$SUBJECT, maxiter=5000)
  # Under the sharp null,create all the hypothesized (Y0 = Y1) 
  Null <- genouts(x, df$treat, 0)
  # returns for each variable a vector with the "fake" distribution relevant to it,under the null.
  return(gendist(Null, perms, prob=probs))
}
```


## Section 3


```{r results='hide'}
# creating a variable that, with the help of the function above, set for each of the 14 variables, a distribution of estimated average treatment effect.
EstDistNull <- sapply(df[vars], fake_distribution)

```

```{r results='asis'}
# display the first "fake" values from the distribution for each variable.
stargazer(head(EstDistNull), type="html")
```


## Section 4


### Sub-section a


```{r}
# setting a TRUE\FALSE matrix such that, a cell in the matrix get the value TRUE, if the "fake" average treatment distribution is higher (in abs) than the observed average treatment distribution. 
Indicator <- apply(EstDistNull, 1, function(x) ifelse(abs(x) >= abs(means[,3]), TRUE, FALSE))

```


### Sub-section b


```{r}
# By counting the number of TRUE appearances in the matrix and dividing it by the total number of permutations,computing the p-value for each variable.
p_values <- apply(Indicator, 1, mean)
```


### Sub-section c


```{r results='asis'}
# display a P Values Table for all the variables.
stargazer(p_values, type = "html", title = "P Values Table", flip = TRUE)

```


## Section 5


According to the P Values, We can infer that the treatment of high quality education setting from infancy to age 5 in poverty areas, did improved IQ levels of the kids throughout the years.
on the other hand, the treatment had no effect what so ever regarding to criminal activities.



